Online Social Cohesion Reflects Real-World Group Action in Syria during the Arab Spring

### 2020-06-01 manuscript.Rmd
Integrating Leah's edits pt. 2

## Social Media and Social Movements

During the Arab Spring, social media provided real-time information about the events transpiring in countries across the Middle East and North Africa. Our understanding about the role of social media in social movements is still evolving. Some argue that these mass movements would not have been possible without
efficient large-scale coordination, both physically on the ground and virtually 
on social media platforms. Along with Facebook, Twitter has been credited as 
essential to the Arab Spring---both for protesters (as a means to coordinate 
with one another and to communicate with the world) and for the government 
response (to monitor and track protestors to crack down on dissent) 
[@comninos2011twitter]. The events of Arab Spring have been thoroughly 
discussed within the political science literature 
[e.g., @lutterbeck2013arab; @koehler2017political], but to our knowledge,
the direct *coupling* between Syrian Twitter cohesion and international action
coordination has not been investigated. Others provide skepticism about the effect of social media on collective action [@Seay_2014].

@@

## Social Cohesion and Social Mobilization
Real-world collective action is known to bring about a sense of togetherness,
also known as “collective effervescence” [@durkheim1995elementary]. This
feeling of being a part of something larger than oneself leads individuals
to act in ways that would not have on their own. Being overcome with this
feeling can lead to violence [@collins2009violence; @spaaij2014sports].
Conversely, increased exposure
to violence can result in positive social cohesion. @gilligan2014civil
found that exposure to fatal wartime violence resulted in increased prosocial
motivations, including cooperative behaviors, feelings of trust, and overall
community-level social cohesion.

>>>>

## Social Media and Social Movements
During the Arab Spring, social media provided real-time information about the events transpiring in countries across the Middle East and North Africa. Our understanding about the role of social media in social movements is still evolving. Some argue that these mass movements would not have been possible without
efficient large-scale coordination, both physically on the ground and virtually 
on social media platforms. Along with Facebook, Twitter has been credited as 
essential to the Arab Spring---both for protesters (as a means to coordinate 
with one another and to communicate with the world) and for the government 
response (to monitor and track protestors to crack down on dissent) 
[@comninos2011twitter]. The events of Arab Spring have been thoroughly 
discussed within the political science literature 
[e.g., @lutterbeck2013arab; @koehler2017political], but to our knowledge,
the direct *coupling* between Syrian Twitter cohesion and international action
coordination has not been investigated. Others provide skepticism about the effect of social media on collective action [@Seay_2014].

@@

## Lexical cohesion: Connectedness of words
The words that comprise social media posts are the first level at which cohesion occurs. In Figure 1, we see four sequential tweets, each containing four words. Each sequential tweet contains some content included in the previous tweets. The linguistic construct of content word overlap measures the relationship across units of meaning – here, tweets – where identical and semantically similar words are repeated [@Graesser_McNamara_Louwerse_Cai_2004]. These repeated lexical items constitute a theme; on Twitter, this might be a particular hashtag, a name, a location, or other nominal category. This localized cohesion is referential; words are repeated and co-referred across tweets, creating a sense of continuity and cohesiveness of meaning and message.

Lexical cohesion also includes latent semantic analysis (LSA) [@Graesser_McNamara_Louwerse_Cai_2004. LSA measures semantic similarity across meaning units using a statistical method called singular value decomposition, an automated method used to detect plagiarism [@Ceska_2008; @Gamallo_Bordag_2011]. In this process, each meaning unit (tweet) is a weighted vector, and the match between the tweets is measured as the cosine between the two vectors. This process provides an empirical relationship between the words used from one tweet to the next. We identify this as lexical cohesion.

The concept of lexical, or semantic, cohesion has already been applied to the analysis of political speech, for example the language of Margaret Thatcher, as well as non-political speech [@Barzilay_Elhadad_1999; @Morris_Hirst_1991]. McNamara and Graesser operationalize the concept of cohesion in their software, Coh-Metrix, an automated tool that analyzes text and discourse  [@Graesser_Jeon_Yan_Cai_2007; @Graesser_McNamara_Louwerse_Cai_2004; [@Graesser_McNamara_Kulikowich_2011]. This software generates measures of co-referential cohesion that captures the amount of overlap in content words (as opposed to function words, like a, the, and and) in adjacent, local, and global sentences in the passage. Content word overlap refers to the proportion of words and word stems that are shared between sentences. A higher proportion of overlap indicates greater cohesion in the sentence.

In estimating cohesion in Twitter posts, tweets that share content words and stems will yield higher cohesion values, even amongst speakers with opposing views on the subject matter. Repetition of terms and concepts is captured by reiteration, which is a type of lexical cohesion in which a subsequent item repeats a previous item, or refers to it through synonymous language [@Halliday_Hasan_1976]. In social media, we can conceive of retweets – reposts or forwards of previous messages in part or in whole – as a form of reiteration that increase message cohesion. Since retweets replicate verbatim some part of another tweet, these tweets would also have higher cohesion. Hashtags function similarly, and are included in our analysis.

```{r figure-1}

# call figure 1 here

## Interlocutor cohesion
Lexical cohesion contributes to the next level of relatedness among tweets: interlocutor cohesion. This is the degree to which people in a conversation, whether in person or online, have aligned their language [@Coupland_Giles_1988]. Known as communication accommodation theory, the degree of alignment between speakers reveals their closeness, status, and deliberative strategy [@Giles_2016]. Speakers can converge and diverge in their communication. Convergence facilitates easier transmission of information, because the speakers – or interlocutors – presume some existing familiarity or knowledge base between them. When you know someone well enough to finish their sentence, this is an example of convergence. Divergence, on the other hand, is more laborious because the speakers must make their presumptions explicit and tacit. Interlocutor cohesion, or convergence, happens when speakers are “singing in the same key” or are “on the same page.” Windsor (2020) describes various types of cohesion, using the example of the Occupy Wall Street movement as a symbol of divergence given the myriad signs, slogans, and interests of the participants [@Windsor_2020]. On the other hand, the “I am a man” protest captured by Ernest Withers during the 1968 santiation workers’ strike in Memphis, TN, or the “Black Lives Matter” placards following the uprising in Ferguson, MO, symbolize convergence and strong interlocutor cohesion [@Edgar_Johnson_2018].

Cognitive and social psychology theories of alignment and coordination lend themselves nicely to the study of social media communication during times of conflict and contentious behavior [@Tollefsen_Dale_Paxton_2013; @Garrod_Pickering_2004; @Pickering_Garrod_2004]. Over the past few decades, this area has garnered significant attention and produced evidence showing that people sometimes linguistically and behaviorally align as they interact more [@Giles_Coupland_Coupland_1991; @Tollefsen_Dale_Paxton_2013]. Similarly, accommodation theory refers to the ways in which individuals and groups linguistically adapt to minimize the social space distance. Speakers in a collective setting pursue strategies which increase their common ground, and this is reflected in the language they use to communicate with one another.

A similar process happens among participants in social media discussions. The process of speakers’ (or tweeters) language becoming more similar is called linguistic style matching (LSM) [@Niederhoffer_Pennebaker_2002]. LSM is a form of synchrony between speakers, akin to linguistic convergence. As Bernieri and Rosenthal (1991) describe, people engaged in linguistic synchrony match each others’ verbal behaviors, adopting not only similar terms, but similar patterns of language [@Bernieri_Rosenthal_1991]. For example, LSM is particularly evident when one speaker introduces swearing into a genteel conversation; the first usage opens the vulgarity floodgates, and subsequent speakers may repeat the initial curse, as well as innovate new curses into the conversation. Fed by lexical cohesion, interlocutor cohesion ebbs and flows along a timeline contextualized by variation in real world events. Interlocutor cohesion may increase preceding an in-person social mobilization event, alerting a repressive government that collective mobilization is happening. Interlocutor cohesion may also respond to exogenous events such as spontaneous, unprovoked militarized conflict. If government forces target a new location using military weapons, this can precipitate online mobilization and discussion of the event.

## Social and Physical Cohesion
In turn, other studies show that linguistic alignment leads to increased social connectedness and solidarity [@Miles_Nind_Macrae_2009], which is associated with increased motivation for participation [@Walton_Cohen_Cwir_Spencer_2012], facilitating successful coordination of group actions toward shared goals [@Fusaroli_Bahrami_Olsen_Roepstorff_Rees_Frith_Tylén_2012; @Richardson_Dale_2005]. This research emphasizes both the importance of and connections between linguistic alignment and behavioral alignment. The more people interact, the more similar their language and linguistic patterns become. This holds true for face-to-face interactions as well as online communication [@Riordan_Dale_Kreuz_Olney_2011}. Given this, we should expect to see increased linguistic cohesion amongst people participating in online discussions about an idea, event, or individual. Given that social media reports in real time about actual events taking place, i.e., protests or demonstrations in which people are assembled together in a location, or attacks that require coordination between members of a unit, the cohesion in social media should be expected to correlate with the social cohesion of the group [@Zeitzoff_2011}.

A socially cohesive group is inherently more threatening to the existing regime because it reflects the professionalization of potential alternate leadership, and collective opposition momentum. Groups with high social cohesion are more difficult to divide and conquer, and more cohesive messages may also garner increased international support. A paper published in Nature argues that global insurgency movements all exhibit a similar pattern to their social structure [@Bohorquez_Gourley_Dixon_Spagat_Johnson_2009}. The authors assert that each violent group has a stable state, and that exogenous events can create perturbations that push the group to become more cohesive, or alternatively to become more fragmented. Rather than frenetic, patchwork messages via Twitter, cohesive messages indicate increased organizational structure, which should incentivize a response from the government. What process is the driving force, though? Does an increase in linguistic cohesion precede or follow increased repression by the government? Do movements get stronger first, and then the government responds, or does the government increase its offensive measures, and thus social groups have a clear opponent around whom they can coalesce both socially and linguistically?

Thus, verbal synchrony leads to physical synchrony. Shared language in an online environment can influence individuals’ propensity to engage in similar coordinated behavior. In the case of the Arab Spring, this may mean social mobilization in public spaces to protest against government brutality. Words in a passage of text – or series of tweets – have some degree of semantic relation. Individuals engaged in a conversation, whether in real life or online, have the tendency to use similar language as they interact, increasing the cohesion of their language. As an extension, multiple individuals collectively communicating can also demonstrate linguistic alignment. In the context of social conflict, the linguistic cohesion of participants in the online conversation should correspond to the degree of social cohesion on the ground. We link these concepts of social and linguistic cohesion together to demonstrate how language serves as an indicator of the degree of social mobilization, using Twitter feeds during the spring of 2012 in Syria.

Real-world collective action is known to bring about a sense of togetherness,
also known as “collective effervescence” [@durkheim1995elementary]. This
feeling of being a part of something larger than oneself leads individuals
to act in ways that would not have on their own. Being overcome with this
feeling can lead to violence [@collins2009violence; @spaaij2014sports].
Conversely, increased exposure
to violence can result in positive social cohesion. @gilligan2014civil
found that exposure to fatal wartime violence resulted in increased prosocial
motivations, including cooperative behaviors, feelings of trust, and overall
community-level social cohesion.

Physical and social cohesion in the world can be modeled using event data which encodes who does what to whom on a given day. This could include events such as the government (GOV) imposing a blockade or fighting with artillery or tanks on a civilian group (CIV). CAMEO (Conflict and Mediation Event Observations) encode a set of activities with numerical values, rescaled to reflect a conflict-cooperation scale [@Schrodt_Hall_2006; @Goldstein_1992]. These event codes include a range of behaviors, from making public statements and appeals, to demanding, threatening, protesting, and using unconventional mass violence. Steinert-Threlkeld (2017) modeled a similar relationship between social media and social mobilization using a Gini coefficient measure, rather than a tf-idf (term frequency-inverse document frequency) measure alongside event data [@Ramos_2003; @Threlkeld_2017]. This pathbreaking work demonstrates that participants on the fringe of a social movement contribute substantially to the overall level of protest. We add to this discussion by demonstrating how sociolinguistic cohesion evolves in response to negative and positive exogenous events.

### Cohesion in Social Movements and Civil Wars
People join social and political resistance movements because they feel a sense of connection to the cause, and because they have been sufficiently convinced to participate. Given that people are generally risk averse, overcoming inertia requires a strong connection to the group identity, or against an individual, event, or political injustice [@Goldstone_1980; Goldstone_2011]. Political resistance offers participants the opportunity to “lose themselves” and relinquish their individual identity in favor of identification with a larger group identity, thus contributing to a cohesive group ready and mobilized for action [@Hoffer_2011. Previous research on social movements supports the idea that participants are likely to join movements when they know and have a personal connection to someone else already involved [@McAdam_1986. The type of collective contentious behavior that manifests in society is a function of the available political opportunity structures. The political opportunity structure, as defined by Kitschelt (1986) are “comprised of specific configuration of resources, institutional arrangements and historical precedents for social mobilization, which facilitate the development of protest movements in some instances and constrain them in others [@Kitschelt_1986].” What emerges as strategies and tactics in contentious politics is in essence, a function of what the system makes possible. This includes the number of people and their disposition, i.e. nonviolent or violent, as well as their internal structure. The more cohesive a group is, the greater threat they pose to the establishment.

Cohesion is also a familiar concept in the study of civil conflicts as it pertains to group affiliation and the ability to overcome collective action problems. Olson’s (Olson, 1965) seminal work on collective action informs the theory of group behavior. Individuals must act in concert – cohesively – to accomplish the task of challenging the authority of the state. Building on Eisinger’s (1973) analysis of social protest movements [@Eisinger_1973, McAdam, Tilly and Tarrow [@Tilly_2004, and Francisco [@Francisco_1995 have provided foundational work on the dynamics of social movements and contentious behavior. More recently, Gates (2002) considers rebel recruitment for participation in civil conflicts as a function of internal cohesion [@Gates_2002]. He notes that rebel movement success is influenced by the degree of group cohesion, which itself is a function of geographic distance, ideology and ethnicity.

There are multiple rebel groups within Syria, some with fighters sourced locally from the disaffected Syrian population, and an increasing number of foreign fighters from transnational terrorist organizations. Gates’ work helps us understand the intergroup competition for participation in the rebel movement, as well as the groups’ abilities to punish defectors for nonparticipation. Foreign fighters would appear to be amply motivated to fight, given their decision to relocate from their home country which is sometimes in the region but other times requires them to travel large distances.

Similarly, Sambanis and Shayo (2013) find in their model that “intense ethnic conflict makes people care more about their ethnic group relative to other groups, and seek to resemble it more [@Sambanis_Shayo_2013].” As ethnic conflict intensifies, groups seeking to become more quintessentially like the typical member of their ethnic group, this should be manifested in their language on social media. Social medial should reflect this delineation amongst ethnic groups and between them. However, this may also include a general identification across ethnic groups in a membership of an opposition to an oppressive leader.

It is this general identification with the larger opposition movement, not the individual identity groups within Syria, which we seek to understand in this paper. It is implausible to expect there to be one cohesive narrative from the multiple groups with disparate incentives. However, they do have important similarities, namely, they are all interested in opposing the current Syrian regime using largely violent means, although there are notable and regular instances of nonviolent protests on Fridays, following prayer services. Thus, what we are able to capture in this study is the degree to which all opposition movements to the Syrian government have a similar content and tone, with a similar ebb and flow. It should be increasingly threatening to the regime if the rebel movement across the disparate groups is growing stronger and more unified, represented by similar language used across groups.

The civil war literature has further explored the role of social cohesion in detail. Collier and Hoeffler identified the role of social homogeneity as having an increased likelihood on social conflict, especially when the society is polarized [@Collier_Hoeffler_2004]. Cederman and Girardin  write that, “In general, nationalist groups derive their cohesion from collective-level identities that resonate deeply with the emotional needs of modern citizens.” Sambanis and Schulhofer-Wohl note that “group cohesion and solidarity (underlie) the offensive advantage (Sambanis & Schulhofer-Wohl, 2009).” Anyanwu (2002) writes, “Another source of rebel military opportunity is social cohesion…an ethnically diverse society may reduce the opportunity for rebellion by limiting the recruitment pool [@Anyanwu_2002].” Referencing Gurr’s work on minorities at risk, Regan and Norton (2005) note that “a religious or linguistic minority might suffer disproportionately in a given society, and this form of aggrievement can lead to unrest across the social lines that distinguish the minority group. Therefore, while collective cohesion is a function of targeted maldistributions, it is individual penury that allows for this collective mobilization [@Gurr_2000; @Regan_Norton_2005].” Thus scholarship concurs that group cohesion matters to group mobilization; we add to this concept by introducing the role of language in facilitating cohesion.

### 2020-06-01 manuscript.tex
Integrating Leah's edits pt. 2

\hypertarget{social-cohesion-and-social-mobilization}{%
\subsection{Social Cohesion and Social Mobilization}\label{social-cohesion-and-social-mobilization}}

@@

>>>>

\hypertarget{lexical-cohesion-connectedness-of-words}{%
\subsection{Lexical cohesion: Connectedness of words}\label{lexical-cohesion-connectedness-of-words}}

The words that comprise social media posts are the first level at which cohesion occurs. In Figure 1, we see four sequential tweets, each containing four words. Each sequential tweet contains some content included in the previous tweets. The linguistic construct of content word overlap measures the relationship across units of meaning -- here, tweets -- where identical and semantically similar words are repeated ({\textbf{???}}). These repeated lexical items constitute a theme; on Twitter, this might be a particular hashtag, a name, a location, or other nominal category. This localized cohesion is referential; words are repeated and co-referred across tweets, creating a sense of continuity and cohesiveness of meaning and message.

Lexical cohesion also includes latent semantic analysis (LSA) {[}({\textbf{???}}). LSA measures semantic similarity across meaning units using a statistical method called singular value decomposition, an automated method used to detect plagiarism ({\textbf{???}}; {\textbf{???}}). In this process, each meaning unit (tweet) is a weighted vector, and the match between the tweets is measured as the cosine between the two vectors. This process provides an empirical relationship between the words used from one tweet to the next. We identify this as lexical cohesion.

The concept of lexical, or semantic, cohesion has already been applied to the analysis of political speech, for example the language of Margaret Thatcher, as well as non-political speech ({\textbf{???}}; {\textbf{???}}). McNamara and Graesser operationalize the concept of cohesion in their software, Coh-Metrix, an automated tool that analyzes text and discourse {[}({\textbf{???}}); ({\textbf{???}}); ({\textbf{???}}). This software generates measures of co-referential cohesion that captures the amount of overlap in content words (as opposed to function words, like a, the, and and) in adjacent, local, and global sentences in the passage. Content word overlap refers to the proportion of words and word stems that are shared between sentences. A higher proportion of overlap indicates greater cohesion in the sentence.

In estimating cohesion in Twitter posts, tweets that share content words and stems will yield higher cohesion values, even amongst speakers with opposing views on the subject matter. Repetition of terms and concepts is captured by reiteration, which is a type of lexical cohesion in which a subsequent item repeats a previous item, or refers to it through synonymous language ({\textbf{???}}). In social media, we can conceive of retweets -- reposts or forwards of previous messages in part or in whole -- as a form of reiteration that increase message cohesion. Since retweets replicate verbatim some part of another tweet, these tweets would also have higher cohesion. Hashtags function similarly, and are included in our analysis.

\hypertarget{interlocutor-cohesion}{%
\subsection{Interlocutor cohesion}\label{interlocutor-cohesion}}

Lexical cohesion contributes to the next level of relatedness among tweets: interlocutor cohesion. This is the degree to which people in a conversation, whether in person or online, have aligned their language ({\textbf{???}}). Known as communication accommodation theory, the degree of alignment between speakers reveals their closeness, status, and deliberative strategy ({\textbf{???}}). Speakers can converge and diverge in their communication. Convergence facilitates easier transmission of information, because the speakers -- or interlocutors -- presume some existing familiarity or knowledge base between them. When you know someone well enough to finish their sentence, this is an example of convergence. Divergence, on the other hand, is more laborious because the speakers must make their presumptions explicit and tacit. Interlocutor cohesion, or convergence, happens when speakers are \enquote{singing in the same key} or are \enquote{on the same page.} Windsor (2020) describes various types of cohesion, using the example of the Occupy Wall Street movement as a symbol of divergence given the myriad signs, slogans, and interests of the participants ({\textbf{???}}). On the other hand, the \enquote{I am a man} protest captured by Ernest Withers during the 1968 santiation workers' strike in Memphis, TN, or the \enquote{Black Lives Matter} placards following the uprising in Ferguson, MO, symbolize convergence and strong interlocutor cohesion ({\textbf{???}}).

Cognitive and social psychology theories of alignment and coordination lend themselves nicely to the study of social media communication during times of conflict and contentious behavior ({\textbf{???}}; {\textbf{???}}; {\textbf{???}}). Over the past few decades, this area has garnered significant attention and produced evidence showing that people sometimes linguistically and behaviorally align as they interact more ({\textbf{???}}; {\textbf{???}}). Similarly, accommodation theory refers to the ways in which individuals and groups linguistically adapt to minimize the social space distance. Speakers in a collective setting pursue strategies which increase their common ground, and this is reflected in the language they use to communicate with one another.

A similar process happens among participants in social media discussions. The process of speakers' (or tweeters) language becoming more similar is called linguistic style matching (LSM) ({\textbf{???}}). LSM is a form of synchrony between speakers, akin to linguistic convergence. As Bernieri and Rosenthal (1991) describe, people engaged in linguistic synchrony match each others' verbal behaviors, adopting not only similar terms, but similar patterns of language ({\textbf{???}}). For example, LSM is particularly evident when one speaker introduces swearing into a genteel conversation; the first usage opens the vulgarity floodgates, and subsequent speakers may repeat the initial curse, as well as innovate new curses into the conversation. Fed by lexical cohesion, interlocutor cohesion ebbs and flows along a timeline contextualized by variation in real world events. Interlocutor cohesion may increase preceding an in-person social mobilization event, alerting a repressive government that collective mobilization is happening. Interlocutor cohesion may also respond to exogenous events such as spontaneous, unprovoked militarized conflict. If government forces target a new location using military weapons, this can precipitate online mobilization and discussion of the event.

\hypertarget{social-and-physical-cohesion}{%
\subsection{Social and Physical Cohesion}\label{social-and-physical-cohesion}}

In turn, other studies show that linguistic alignment leads to increased social connectedness and solidarity ({\textbf{???}}), which is associated with increased motivation for participation ({\textbf{???}}), facilitating successful coordination of group actions toward shared goals ({\textbf{???}}; {\textbf{???}}). This research emphasizes both the importance of and connections between linguistic alignment and behavioral alignment. The more people interact, the more similar their language and linguistic patterns become. This holds true for face-to-face interactions as well as online communication {[}({\textbf{???}})\}. Given this, we should expect to see increased linguistic cohesion amongst people participating in online discussions about an idea, event, or individual. Given that social media reports in real time about actual events taking place, i.e., protests or demonstrations in which people are assembled together in a location, or attacks that require coordination between members of a unit, the cohesion in social media should be expected to correlate with the social cohesion of the group {[}({\textbf{???}})\}.

A socially cohesive group is inherently more threatening to the existing regime because it reflects the professionalization of potential alternate leadership, and collective opposition momentum. Groups with high social cohesion are more difficult to divide and conquer, and more cohesive messages may also garner increased international support. A paper published in Nature argues that global insurgency movements all exhibit a similar pattern to their social structure {[}({\textbf{???}})\}. The authors assert that each violent group has a stable state, and that exogenous events can create perturbations that push the group to become more cohesive, or alternatively to become more fragmented. Rather than frenetic, patchwork messages via Twitter, cohesive messages indicate increased organizational structure, which should incentivize a response from the government. What process is the driving force, though? Does an increase in linguistic cohesion precede or follow increased repression by the government? Do movements get stronger first, and then the government responds, or does the government increase its offensive measures, and thus social groups have a clear opponent around whom they can coalesce both socially and linguistically?

Thus, verbal synchrony leads to physical synchrony. Shared language in an online environment can influence individuals' propensity to engage in similar coordinated behavior. In the case of the Arab Spring, this may mean social mobilization in public spaces to protest against government brutality. Words in a passage of text -- or series of tweets -- have some degree of semantic relation. Individuals engaged in a conversation, whether in real life or online, have the tendency to use similar language as they interact, increasing the cohesion of their language. As an extension, multiple individuals collectively communicating can also demonstrate linguistic alignment. In the context of social conflict, the linguistic cohesion of participants in the online conversation should correspond to the degree of social cohesion on the ground. We link these concepts of social and linguistic cohesion together to demonstrate how language serves as an indicator of the degree of social mobilization, using Twitter feeds during the spring of 2012 in Syria.

@@

Physical and social cohesion in the world can be modeled using event data which encodes who does what to whom on a given day. This could include events such as the government (GOV) imposing a blockade or fighting with artillery or tanks on a civilian group (CIV). CAMEO (Conflict and Mediation Event Observations) encode a set of activities with numerical values, rescaled to reflect a conflict-cooperation scale ({\textbf{???}}; {\textbf{???}}). These event codes include a range of behaviors, from making public statements and appeals, to demanding, threatening, protesting, and using unconventional mass violence. Steinert-Threlkeld (2017) modeled a similar relationship between social media and social mobilization using a Gini coefficient measure, rather than a tf-idf (term frequency-inverse document frequency) measure alongside event data ({\textbf{???}}; {\textbf{???}}). This pathbreaking work demonstrates that participants on the fringe of a social movement contribute substantially to the overall level of protest. We add to this discussion by demonstrating how sociolinguistic cohesion evolves in response to negative and positive exogenous events.

\hypertarget{cohesion-in-social-movements-and-civil-wars}{%
\subsubsection{Cohesion in Social Movements and Civil Wars}\label{cohesion-in-social-movements-and-civil-wars}}

People join social and political resistance movements because they feel a sense of connection to the cause, and because they have been sufficiently convinced to participate. Given that people are generally risk averse, overcoming inertia requires a strong connection to the group identity, or against an individual, event, or political injustice {[}({\textbf{???}}); Goldstone\_2011{]}. Political resistance offers participants the opportunity to \enquote{lose themselves} and relinquish their individual identity in favor of identification with a larger group identity, thus contributing to a cohesive group ready and mobilized for action {[}({\textbf{???}}). Previous research on social movements supports the idea that participants are likely to join movements when they know and have a personal connection to someone else already involved {[}({\textbf{???}}). The type of collective contentious behavior that manifests in society is a function of the available political opportunity structures. The political opportunity structure, as defined by Kitschelt (1986) are \enquote{comprised of specific configuration of resources, institutional arrangements and historical precedents for social mobilization, which facilitate the development of protest movements in some instances and constrain them in others ({\textbf{???}}).} What emerges as strategies and tactics in contentious politics is in essence, a function of what the system makes possible. This includes the number of people and their disposition, i.e.~nonviolent or violent, as well as their internal structure. The more cohesive a group is, the greater threat they pose to the establishment.

Cohesion is also a familiar concept in the study of civil conflicts as it pertains to group affiliation and the ability to overcome collective action problems. Olson's (Olson, 1965) seminal work on collective action informs the theory of group behavior. Individuals must act in concert -- cohesively -- to accomplish the task of challenging the authority of the state. Building on Eisinger's (1973) analysis of social protest movements {[}({\textbf{???}}), McAdam, Tilly and Tarrow {[}({\textbf{???}}), and Francisco {[}({\textbf{???}}) have provided foundational work on the dynamics of social movements and contentious behavior. More recently, Gates (2002) considers rebel recruitment for participation in civil conflicts as a function of internal cohesion ({\textbf{???}}). He notes that rebel movement success is influenced by the degree of group cohesion, which itself is a function of geographic distance, ideology and ethnicity.

There are multiple rebel groups within Syria, some with fighters sourced locally from the disaffected Syrian population, and an increasing number of foreign fighters from transnational terrorist organizations. Gates' work helps us understand the intergroup competition for participation in the rebel movement, as well as the groups' abilities to punish defectors for nonparticipation. Foreign fighters would appear to be amply motivated to fight, given their decision to relocate from their home country which is sometimes in the region but other times requires them to travel large distances.

Similarly, Sambanis and Shayo (2013) find in their model that \enquote{intense ethnic conflict makes people care more about their ethnic group relative to other groups, and seek to resemble it more ({\textbf{???}}).} As ethnic conflict intensifies, groups seeking to become more quintessentially like the typical member of their ethnic group, this should be manifested in their language on social media. Social medial should reflect this delineation amongst ethnic groups and between them. However, this may also include a general identification across ethnic groups in a membership of an opposition to an oppressive leader.

It is this general identification with the larger opposition movement, not the individual identity groups within Syria, which we seek to understand in this paper. It is implausible to expect there to be one cohesive narrative from the multiple groups with disparate incentives. However, they do have important similarities, namely, they are all interested in opposing the current Syrian regime using largely violent means, although there are notable and regular instances of nonviolent protests on Fridays, following prayer services. Thus, what we are able to capture in this study is the degree to which all opposition movements to the Syrian government have a similar content and tone, with a similar ebb and flow. It should be increasingly threatening to the regime if the rebel movement across the disparate groups is growing stronger and more unified, represented by similar language used across groups.

The civil war literature has further explored the role of social cohesion in detail. Collier and Hoeffler identified the role of social homogeneity as having an increased likelihood on social conflict, especially when the society is polarized ({\textbf{???}}). Cederman and Girardin write that, \enquote{In general, nationalist groups derive their cohesion from collective-level identities that resonate deeply with the emotional needs of modern citizens.} Sambanis and Schulhofer-Wohl note that \enquote{group cohesion and solidarity (underlie) the offensive advantage (Sambanis \& Schulhofer-Wohl, 2009).} Anyanwu (2002) writes, \enquote{Another source of rebel military opportunity is social cohesion\ldots{}an ethnically diverse society may reduce the opportunity for rebellion by limiting the recruitment pool ({\textbf{???}}).} Referencing Gurr's work on minorities at risk, Regan and Norton (2005) note that \enquote{a religious or linguistic minority might suffer disproportionately in a given society, and this form of aggrievement can lead to unrest across the social lines that distinguish the minority group. Therefore, while collective cohesion is a function of targeted maldistributions, it is individual penury that allows for this collective mobilization ({\textbf{???}}; {\textbf{???}}).} Thus scholarship concurs that group cohesion matters to group mobilization; we add to this concept by introducing the role of language in facilitating cohesion.
